---
title: "Nowcasting the current size of the COVID-19 outbreak in the United States"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}

library(tidyverse)
library(tibbletime)
library(padr)
library(tvReg)
library(forecast)
source('R/package.R')
source('R/deconcolve.R')
source('R/simple_tdar.R')
source('R/package_tv.R')
source('R/nowcast.R')

```

## {.tabset .tabset-fade}

### Algorithm

**At any given time, most COVID-19 cases are circulating in the community and not known to us.**

We wish to estimate the **total current size of the COVID-19 outbreak**, i.e. the total number of unnotified individuals currently infected with SARS-CoV2. We need to estimate this number from **case reports** (the number of newly confirmed cases on a given day).

Case reports do not include infected persons who are symptomatic but not yet diagnosed, nor infected persons with latent infections (people who cary the virus but are not *yet* symptomatic). In addition, a number of individuals may be infected but never develop sysmptoms. It is possible that these these *asymptomatic infections* may still transmit the virus.

To estimate the size of the outbreak at any given time, we first estimate the number of symptomatic and latent infections in the community, accounting for cases that would be missed because they would never show symptoms.

The infection fatality rate of COVID-19 is known with a reasonable degree of precision to be around 0.9%.
We reasoned that fatal cases are much less likely to be overlooked than non-fatal cases (particularly asymptomatic cases and cases with mild symtoms).
Additionally, the distribution of time from onset of symptoms to death has been estimated.
Taken together, these pieces of information allowthe construction of a symptom onset curve in a manner analogous to that used for case notification data.
The ratio of the two different estimates for the number of persons acquiring infection on any given day is an estimate of the **ascertainment rate**.

**Details:**

From case notification reports (the number of newly confirmed cases), we use Monte Carlo simulation and estimated probability distributions to construct a synthetic **line list** of symptom onset times for all cases that are eventually notified. The average **delay from symtpom onset to case notification** for the US is 2.67 days^[Analysis by the Center for the Ecology of Infectious Diseases of US data. See supplemental information.]. Simulated symtom onset times are assumed to be exponentially distributed with a mean of 2.67 days before the reporting date.

A second simulation generates synthetic exposure dates for each infected individual. The average **incubation period** for the US is 7.67 days^[Analysis by the Center for the Ecology of Infectious Diseases of US data suggests a mean incubation period of 7.67 days. See supplemental information.]. Individual simulated exposure times are distributed according to a gamma distribution with a mean of 7.67 days before symptom onset (gshape parameter: 4.18).

Now we have a synthetic line list that generates the observed number of case notifications  and includes exposure date, symptom onset date, and notification date.

From the line list, we tally the **number of individuals with latent infections** on each date to produce a curve of the numbr of latent cases in the community. We do the same to find the **number of individuals symptomatic infections** on each date. These curves represent only individuals with latent and symptomatic cases that were eventually reported.

Based on testing of passengers on the Princes Cruises ship in Yokohama, Japan, Mizumoto *et al.* estimated that the **proportion of cases that are asymptomatic** is 34.6% (95% CrI: 29.4%–39.8%)^[Kenji Mizumoto, Katsushi Kagaya, Alexander Zarebski, Gerardo Chowell. Estimating the Asymptomatic Ratio of 2019 Novel Coronavirus onboard the Princess Cruises Ship, 2020. medRxiv preprint. <https://doi.org/10.1101/2020.02.20.20025866>]. We assume asymptomatic cases are never detected. To account for these, we multiply both the number of symptomatic cases and the number of latent cases by $\frac{1}{34.6\%}$. 

Since the symptom onsets occur on average 2.67 days before reporting, we are only sure to have an accurate estimate for the number of symptomatic individuals at some time prior to 2.67 days befor the present. We therefore discard estimates for the last several days, and use statistical forecasting to extrapolate from the observed trajectory of symptomatic cases to the present time. This is the "nowcast" of symptomatic individuals. We repeat the process for latent cases, replacing values for times within 7.67 dyas of the present time by a forecasted estimate.

The sum of the latent and symptomatic cases at any time provides an estimate of number of asymptomatic and unnotified cases.

<!-- **Natural infectious period** $\frac{1}{\gamma_0}$ -->

<!-- The natural infectious period $\frac{1}{\gamma_0}$ in the absence of containment is assumed to be 7 days. -->
<!-- Set params for US -->
```{r params, echo=FALSE, message = FALSE, warning = FALSE, cache = TRUE}
US.params <- list(admin = "US", # state or country
                    q = 1, # Detection probability
                    a = 0.346, # Proportion of cases that are asymptomatic
                    c = 1, # Transmissability of undetectable cases
                    effective.infectious.period = list(dist="exponential", mean=2.67),
                    infectious.period = list(dist="exponential", mean=7),
                    incubation.period = list(dist="gamma", mean=7.67, shape=4.1775942)
                    )
```


```{r params, echo=FALSE, message = FALSE, warning = FALSE, cache = TRUE}
US.params <- list(admin = "US", # state or country
                    q = 1, # Detection probability
                    a = 0.346, # Proportion of cases that are asymptomatic
                    c = 1, # Transmissability of undetectable cases
                    effective.infectious.period = list(dist="exponential", mean=2.67),
                    infectious.period = list(dist="exponential", mean=7),
                    incubation.period = list(dist="gamma", mean=7.67, shape=4.1775942)
                    )
```


### Data

**US Case Reports by State**

US data are maintained by [CEID](http://ceid.uga.edu) and available  [here](https://docs.google.com/spreadsheets/d/10NYNf4UX7R_nNdM-jkbo7A5SGkB8sUVMIiqU7e_-miU/edit#gid=0).

Data are compiled from <https://en.wikipedia.org/wiki/2020_coronavirus_pandemic_in_the_United_States>, which is updated by anonymous contributors.

We aim to update our dataset daily.

```{r data, include=FALSE, message = FALSE, warning = FALSE, cache = FALSE}
# read data directly to google sheet.

sheet <- "https://docs.google.com/spreadsheets/d/10NYNf4UX7R_nNdM-jkbo7A5SGkB8sUVMIiqU7e_-miU/edit#gid=0"

US <- googlesheets4::sheets_read(sheet)

names(US)[1] <- "Date"

US <- US %>% 
  rename_all(list(~make.names(.))) %>% 
  mutate_if(is.list, fixcolumns) %>% 
  mutate(Date = as.Date(Date)) %>% 
  padr::pad(start_val = as.Date("2019-12-01")) %>%
  replace(., is.na(.), 0) %>%
  tibbletime::as_tbl_time(index=Date) %>% 
  mutate(US = rowSums(.[,-1],na.rm = TRUE))

```

<!-- calculate nowcasts for US states -->
```{r nowcasts, include=FALSE, echo=FALSE, message = FALSE, warning = FALSE, cache = TRUE}

get_US_nowcast <- function(state) {
  US %>% select(Date, cases = state) %>% 
    tbl_time(index = Date) %>% nowcast_from_case_reports(US.params)
}

US.all <- get_US_nowcast("US")
AK <- get_US_nowcast("AK")
AL <- get_US_nowcast("AL")
AR <- get_US_nowcast("AR")
AZ <- get_US_nowcast("AZ")
CA <- get_US_nowcast("CA")
CO <- get_US_nowcast("CO")
CT <- get_US_nowcast("CT")
DC <- get_US_nowcast("DC")
DE <- get_US_nowcast("DE")
FL <- get_US_nowcast("FL")
GA <- get_US_nowcast("GA")
HI <- get_US_nowcast("HI")
IA <- get_US_nowcast("IA")
IL <- get_US_nowcast("IL")
IN <- get_US_nowcast("IN")
KS <- get_US_nowcast("KS")
KY <- get_US_nowcast("KY")
LA <- get_US_nowcast("LA")
MA <- get_US_nowcast("MA")
MD <- get_US_nowcast("MD")
ME <- get_US_nowcast("ME")
MI <- get_US_nowcast("MI")
MN <- get_US_nowcast("MN")
MO <- get_US_nowcast("MO")
MS <- get_US_nowcast("MS")
MT <- get_US_nowcast("MT")
NC <- get_US_nowcast("NC")
ND <- get_US_nowcast("ND")
NE <- get_US_nowcast("NE")
NH <- get_US_nowcast("NH")
NJ <- get_US_nowcast("NJ")
NM <- get_US_nowcast("NM")
NV <- get_US_nowcast("NV")
NY <- get_US_nowcast("NY")
OH <- get_US_nowcast("OH")
OK <- get_US_nowcast("OK")
OR <- get_US_nowcast("OR")
PA <- get_US_nowcast("PA")
RI <- get_US_nowcast("RI")
SC <- get_US_nowcast("SC")
SD <- get_US_nowcast("SD")
TN <- get_US_nowcast("TN")
TX <- get_US_nowcast("TX")
UT <- get_US_nowcast("UT")
VA <- get_US_nowcast("VA")
VT <- get_US_nowcast("VT")
WA <- get_US_nowcast("WA")
WI <- get_US_nowcast("WI")
WY <- get_US_nowcast("WY")
```

### USA
Nowcast (outbreak size) = E + I = the number of non-isolated infected individuals
```{r US, echo=FALSE, message = FALSE, warning = FALSE, cache = FALSE}
plot_nowcast_from_case_reports(US.all)
```

### AK
Nowcast (outbreak size) = E + I = the number of non-isolated infected individuals
```{r AK, echo=FALSE, message = FALSE, warning = FALSE, cache = FALSE}
plot_nowcast_from_case_reports(AK)
```

### AL
Nowcast (outbreak size) = E + I = the number of non-isolated infected individuals
```{r AL, echo=FALSE, message = FALSE, warning = FALSE, cache = FALSE}
plot_nowcast_from_case_reports(AL)
```

### AR
Nowcast (outbreak size) = E + I = the number of non-isolated infected individuals
```{r AR, echo=FALSE, message = FALSE, warning = FALSE, cache = FALSE}
plot_nowcast_from_case_reports(AR)
```

### AZ
Nowcast (outbreak size) = E + I = the number of non-isolated infected individuals
```{r AZ, echo=FALSE, message = FALSE, warning = FALSE, cache = FALSE}
plot_nowcast_from_case_reports(AZ)
```

### CA
Nowcast (outbreak size) = E + I = the number of non-isolated infected individuals
```{r CA, echo=FALSE, message = FALSE, warning = FALSE, cache = FALSE}
plot_nowcast_from_case_reports(CA)
```

### CO
Nowcast (outbreak size) = E + I = the number of non-isolated infected individuals
```{r CO, echo=FALSE, message = FALSE, warning = FALSE, cache = FALSE}
plot_nowcast_from_case_reports(CO)
```

### CT
Nowcast (outbreak size) = E + I = the number of non-isolated infected individuals
```{r CT, echo=FALSE, message = FALSE, warning = FALSE, cache = FALSE}
plot_nowcast_from_case_reports(CT)
```

### DC
Nowcast (outbreak size) = E + I = the number of non-isolated infected individuals
```{r DC, echo=FALSE, message = FALSE, warning = FALSE, cache = FALSE}
plot_nowcast_from_case_reports(DC)
```

### DE
Nowcast (outbreak size) = E + I = the number of non-isolated infected individuals
```{r DE, echo=FALSE, message = FALSE, warning = FALSE, cache = FALSE}
plot_nowcast_from_case_reports(DE)
```

### FL
Nowcast (outbreak size) = E + I = the number of non-isolated infected individuals
```{r FL, echo=FALSE, message = FALSE, warning = FALSE, cache = FALSE}
plot_nowcast_from_case_reports(FL)
```

### GA
Nowcast (outbreak size) = E + I = the number of non-isolated infected individuals
```{r GA, echo=FALSE, message = FALSE, warning = FALSE, cache = FALSE}
plot_nowcast_from_case_reports(GA)
```

### HI
Nowcast (outbreak size) = E + I = the number of non-isolated infected individuals
```{r HI, echo=FALSE, message = FALSE, warning = FALSE, cache = FALSE}
plot_nowcast_from_case_reports(HI)
```

### IA
Nowcast (outbreak size) = E + I = the number of non-isolated infected individuals
```{r IA, echo=FALSE, message = FALSE, warning = FALSE, cache = FALSE}
plot_nowcast_from_case_reports(IA)
```

### IL
Nowcast (outbreak size) = E + I = the number of non-isolated infected individuals
```{r IL, echo=FALSE, message = FALSE, warning = FALSE, cache = FALSE}
plot_nowcast_from_case_reports(IL)
```

### IN
Nowcast (outbreak size) = E + I = the number of non-isolated infected individuals
```{r IN, echo=FALSE, message = FALSE, warning = FALSE, cache = FALSE}
plot_nowcast_from_case_reports(IN)
```

### KS
Nowcast (outbreak size) = E + I = the number of non-isolated infected individuals
```{r KS, echo=FALSE, message = FALSE, warning = FALSE, cache = FALSE}
plot_nowcast_from_case_reports(KS)
```

### KY
Nowcast (outbreak size) = E + I = the number of non-isolated infected individuals
```{r KY, echo=FALSE, message = FALSE, warning = FALSE, cache = FALSE}
plot_nowcast_from_case_reports(KY)
```

### LA
Nowcast (outbreak size) = E + I = the number of non-isolated infected individuals
```{r LA, echo=FALSE, message = FALSE, warning = FALSE, cache = FALSE}
plot_nowcast_from_case_reports(LA)
```

### MA
Nowcast (outbreak size) = E + I = the number of non-isolated infected individuals
```{r MA, echo=FALSE, message = FALSE, warning = FALSE, cache = FALSE}
plot_nowcast_from_case_reports(MA)
```

### MD
Nowcast (outbreak size) = E + I = the number of non-isolated infected individuals
```{r MD, echo=FALSE, message = FALSE, warning = FALSE, cache = FALSE}
plot_nowcast_from_case_reports(MD)
```

### ME
Nowcast (outbreak size) = E + I = the number of non-isolated infected individuals
```{r ME, echo=FALSE, message = FALSE, warning = FALSE, cache = FALSE}
plot_nowcast_from_case_reports(ME)
```

### MI
Nowcast (outbreak size) = E + I = the number of non-isolated infected individuals
```{r MI, echo=FALSE, message = FALSE, warning = FALSE, cache = FALSE}
plot_nowcast_from_case_reports(MI)
```

### MN
Nowcast (outbreak size) = E + I = the number of non-isolated infected individuals
```{r MN, echo=FALSE, message = FALSE, warning = FALSE, cache = FALSE}
plot_nowcast_from_case_reports(MN)
```

### MO
Nowcast (outbreak size) = E + I = the number of non-isolated infected individuals
```{r MO, echo=FALSE, message = FALSE, warning = FALSE, cache = FALSE}
plot_nowcast_from_case_reports(MO)
```

### MS
Nowcast (outbreak size) = E + I = the number of non-isolated infected individuals
```{r MS, echo=FALSE, message = FALSE, warning = FALSE, cache = FALSE}
plot_nowcast_from_case_reports(MS)
```

### MT
Nowcast (outbreak size) = E + I = the number of non-isolated infected individuals
```{r MT, echo=FALSE, message = FALSE, warning = FALSE, cache = FALSE}
plot_nowcast_from_case_reports(MT)
```

### NC
Nowcast (outbreak size) = E + I = the number of non-isolated infected individuals
```{r NC, echo=FALSE, message = FALSE, warning = FALSE, cache = FALSE}
plot_nowcast_from_case_reports(NC)
```

### ND
Nowcast (outbreak size) = E + I = the number of non-isolated infected individuals
```{r ND, echo=FALSE, message = FALSE, warning = FALSE, cache = FALSE}
plot_nowcast_from_case_reports(ND)
```

### NE
Nowcast (outbreak size) = E + I = the number of non-isolated infected individuals
```{r NE, echo=FALSE, message = FALSE, warning = FALSE, cache = FALSE}
plot_nowcast_from_case_reports(NE)
```

### NH
Nowcast (outbreak size) = E + I = the number of non-isolated infected individuals
```{r NH, echo=FALSE, message = FALSE, warning = FALSE, cache = FALSE}
plot_nowcast_from_case_reports(NH)
```

### NJ
Nowcast (outbreak size) = E + I = the number of non-isolated infected individuals
```{r NJ, echo=FALSE, message = FALSE, warning = FALSE, cache = FALSE}
plot_nowcast_from_case_reports(NJ)
```

### NM
Nowcast (outbreak size) = E + I = the number of non-isolated infected individuals
```{r NM, echo=FALSE, message = FALSE, warning = FALSE, cache = FALSE}
plot_nowcast_from_case_reports(NM)
```

### NV
Nowcast (outbreak size) = E + I = the number of non-isolated infected individuals
```{r NV, echo=FALSE, message = FALSE, warning = FALSE, cache = FALSE}
plot_nowcast_from_case_reports(NV)
```

### NY
Nowcast (outbreak size) = E + I = the number of non-isolated infected individuals
```{r NY, echo=FALSE, message = FALSE, warning = FALSE, cache = FALSE}
plot_nowcast_from_case_reports(NY)
```

### OH
Nowcast (outbreak size) = E + I = the number of non-isolated infected individuals
```{r OH, echo=FALSE, message = FALSE, warning = FALSE, cache = FALSE}
plot_nowcast_from_case_reports(OH)
```

### OK
Nowcast (outbreak size) = E + I = the number of non-isolated infected individuals
```{r OK, echo=FALSE, message = FALSE, warning = FALSE, cache = FALSE}
plot_nowcast_from_case_reports(OK)
```

### OR
Nowcast (outbreak size) = E + I = the number of non-isolated infected individuals
```{r OR, echo=FALSE, message = FALSE, warning = FALSE, cache = FALSE}
plot_nowcast_from_case_reports(OR)
```

### PA
Nowcast (outbreak size) = E + I = the number of non-isolated infected individuals
```{r PA, echo=FALSE, message = FALSE, warning = FALSE, cache = FALSE}
plot_nowcast_from_case_reports(PA)
```

### RI
Nowcast (outbreak size) = E + I = the number of non-isolated infected individuals
```{r RI, echo=FALSE, message = FALSE, warning = FALSE, cache = FALSE}
plot_nowcast_from_case_reports(RI)
```

### SC
Nowcast (outbreak size) = E + I = the number of non-isolated infected individuals
```{r SC, echo=FALSE, message = FALSE, warning = FALSE, cache = FALSE}
plot_nowcast_from_case_reports(SC)
```

### SD
Nowcast (outbreak size) = E + I = the number of non-isolated infected individuals
```{r SD, echo=FALSE, message = FALSE, warning = FALSE, cache = FALSE}
plot_nowcast_from_case_reports(SD)
```

### TN
Nowcast (outbreak size) = E + I = the number of non-isolated infected individuals
```{r TN, echo=FALSE, message = FALSE, warning = FALSE, cache = FALSE}
plot_nowcast_from_case_reports(TN)
```

### TX
Nowcast (outbreak size) = E + I = the number of non-isolated infected individuals
```{r TX, echo=FALSE, message = FALSE, warning = FALSE, cache = FALSE}
plot_nowcast_from_case_reports(TX)
```

### UT
Nowcast (outbreak size) = E + I = the number of non-isolated infected individuals
```{r UT, echo=FALSE, message = FALSE, warning = FALSE, cache = FALSE}
plot_nowcast_from_case_reports(UT)
```

### VA
Nowcast (outbreak size) = E + I = the number of non-isolated infected individuals
```{r VA, echo=FALSE, message = FALSE, warning = FALSE, cache = FALSE}
plot_nowcast_from_case_reports(VA)
```

### VT
Nowcast (outbreak size) = E + I = the number of non-isolated infected individuals
```{r VT, echo=FALSE, message = FALSE, warning = FALSE, cache = FALSE}
plot_nowcast_from_case_reports(VT)
```

### WA
Nowcast (outbreak size) = E + I = the number of non-isolated infected individuals
```{r WA, echo=FALSE, message = FALSE, warning = FALSE, cache = FALSE}
plot_nowcast_from_case_reports(WA)
```

### WI
Nowcast (outbreak size) = E + I = the number of non-isolated infected individuals
```{r WI, echo=FALSE, message = FALSE, warning = FALSE, cache = FALSE}
plot_nowcast_from_case_reports(WI)
```

### WY
Nowcast (outbreak size) = E + I = the number of non-isolated infected individuals
```{r WY, echo=FALSE, message = FALSE, warning = FALSE, cache = FALSE}
plot_nowcast_from_case_reports(WY)
```

