---
title: "Nowcasting the current size of the COVID-19 outbreak in China"
author: "Tim Wildauer, Rachel Mercaldo, Eric Marty, John Drake"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}

library(tidyverse)
library(tibbletime)
library(padr)
library(tvReg)
library(forecast)
source('R/package.R')
source('R/deconcolve.R')
source('R/simple_tdar.R')
source('R/package_tv.R')

# functions

# function to trim NA's from a vector
trim_na <- function(x) {x[!is.na(x)]}

# function to pad a vector with NA's
pad_na <- function(x,n,front=FALSE) {
  if(front==TRUE){
    x <- c(rep.int(NA,n),x)
  }else{
    length(x) <- length(x)+n
  }
  return(x)
}

# function to replace tail of vector or dataframe with NA's
tail_na <- function(data, tail, columns=NULL) {
  d <- data
  if (is.vector(d)) {
    n <- length(d)
    # replace(data, (n-tail):n, NA)
    d[(n-tail[1]+1):n] <- NA
  }else{
    if (is.data.frame(d)) {
      n <- nrow(d)
      if(!is.null(columns)){c <- columns}else{c <- 1:length(data)}
      d[(n-tail[1]+1):n,c] <- NA
    }
  }
  return(d)
}

# function to calculate confidence intervals across rows of a dataframe
rowwise_confidence_intervals <- function(df,interval) {
  out <- list()
  for (i in 1:nrow(df)) {
    v <- unlist(df[i,])
    v_sd <- sd(v)
    v_n <- length(v)
    v_mean <- mean(v)
    error <- qt((interval + 1)/2, df = v_n - 1) * v_sd / sqrt(v_n)
    out$lower[i] <- v_mean - error
    out$upper[i] <- v_mean + error
  }
  return(out)
}

```


# Problem

We wish to estimate the total current size of the COVID-19 outbreak, i.e. the total number of people who are currently infected with SARS-CoV2.  

We need to estimate this number from the current count of confirmed cases of COVID-19. The count of confirmed cases does not include those who are symptomatic but whose cases have not been confirmed, those who have been exposed and cary the virus but are not yet symptomatic, and asymptomatic cases (those who are exposed but never develop symptoms).

In order to estimate outbreak size, we assume a model that supposes that at each point in time $t$, every individual in the population may be classified according to one of the following mutually exclusive segments, in accordance with the modified SEIR stochastic model used in <http://2019-coronavirus-tracker.com/stochastic.html>:

$S(t)$ Susceptible individuals
$E(t)$ Exposed individuals in the community (Individuals with a latent infection, not yet able to transmit)
$I(t)$ Infectious individuals in the community
$H(t)$ Individuals with newly confirmed cases (hospitalized or otherwise removed from the community but still infectious)
$R(t)$ Discharged (and no longer infectious) individuals

The total oubreak size $Q(t) = E(t) + I(t) + H(t)$

Here, we lay out a procedure to estimate these time varying populations from a time series of confirmed cases $C(t)$ using deconvolution-based backcasting and time series forecasting. We apply our method to the outbreak in China.

## Model details

We are given only a time series of new case reports $C(t)$. Since we consider that case reporting occurs immediately upon hospitalization, we assume a simplified SEIR model with no $H$ class, and the total outbreak size $Q(t) = E(t) + I(t)$. (If in other situations it is determined that there is a lag between hospitalization and case reporting, as may be the case early in an outbreak, then we could reintroduce the $H$ class.)

**Detection probability** $q(t)$

Following [J.M. Drake & P. Rohani. A stochastic model for the transmission of 2019-nCov in Wuhan](http://2019-coronavirus-tracker.com/stochastic-model.html), we assume the case detection probability to be time varying, with $q(t)$ assumed to have been quite low before the opening of fever clinics on 9 January, 2020, and to have stepped up after January 9. Drake and Rohani have roughly estimated the baseline case detection rate to be $q_0 = 0.11$, and the post January 9 rate to be $q_1 = 0.11$.

```{r}
q_fcn <- function(date) {if_else(date < '2020-01-10', 0.11, 0.98)}
```


**Proportion of cases that are asymptomatic** $a = 0.346$

Based on testing of passengers on the Princes Cruises ship in Yokohama, Japan, Mizumoto *et al.* estimated that the percentage of cases that are asymptomatic $a$ is 34.6% (95% CrI: 29.4%–39.8%) [^1]. 

[^1] : Kenji Mizumoto, Katsushi Kagaya, Alexander Zarebski, Gerardo Chowell. Estimating the Asymptomatic Ratio of 2019 Novel Coronavirus onboard the Princess Cruises Ship, 2020. medRxiv preprint. <https://doi.org/10.1101/2020.02.20.20025866>

```{r}
a_fcn <- function() {0.346}
```

**Transmissability of undetectable cases** $c = 0.52$

The coefficient of relative transmissibility $c$ represents the transmissibility of undetected cases relative to detected cases. Shaman *et al.* ()[^2] have estimated that the transmissibility in undetected cases is approximately 52% of that in detected cases. Therefore we set $c = 0.52$.

[^2] : Shaman *et al.*, personal communication with John Drake

```{r}
c_fcn <- function() {0.52}
```

**Detectable and undetectable cases**

Exposed cases are either detectable $E_d$ or detectable $E_u$ according to $q(t)$ and $a$. Note, detection parameters $q(t)$ and $a$ are applied not to the case reports, but to the exposed and infectious classes, considering that a certain percentage of these are "detectable" (i.e. eventually detected and represented in the times series of case reports $C(t)$. So $E_u(t) = q(t)E_d(t) $ and $E(t) = E_u(t) + E_d(t)$.

Infectious cases are also detected $I_d$ or undetected $I_u$, and $I(t) = I_u(t) + I_d(t)$. Undetected infectious cases depend on the detection probability $q(t)$, proportion asymptomatic $a$ and coefficient of relative transmissibility $c$.


**Infectious period**

The period between symptom onset and hospitalization $\frac{1}{\gamma}$ is time dependent and affected by containment efforts. In the simplified SEIR model, $\frac{1}{\gamma}$ is the period between symptom onset and case report. 

The natural infectious period $\frac{1}{\gamma_0}$ in the absence of containment is assumed to be 7 days.

```{r}
mean.infectious.period <- 7
gamma0_fcn <- function() {1/mean.infectious.period}
```

**Incubation period**

Becker has estimated the period between exposure and symptom onset $\frac{1}{\sigma}$ to be 6.4 days[^2].

[^2] : Becker (citation needed)

`mean.incubation.period <- 6.4`

Need a mean and shape parameter. Tim uses:

`incubation.shape <- 6.2`
`incubation.mean <- 5.6`

```{r}
incubation.shape <- 6.2
incubation.mean <- 5.6
incubation.scale <- incubation.mean/incubation.shape
incubation.sd <- sqrt(incubation.shape*incubation.scale^2)
distribution <- "gamma" # not sure where this gets used. LOOK INSIDE FUNCTION FOR USE OF GLOBAL VARIABLE

sigma_fcn <- function() {1/incubation.mean}
```


## Procedure

Given the time series of new case reports $C(t)$, we first derive a time series of newly symptomatic individuals $I_s(t)$ by deconvolution. From $I_s(t)$ and the natural infectious period $\frac{1}{\gamma_0}$ we build a time series of infectious (and eventually detected) individuals $I_d(t)$.

We derive $I_u(t)$ from $I_d(t)$ and the detection parameters $q(t)$ and $a$ and the transmissibility parameter $c$. 

The time series of exposed $E(t)$ is derived by a second deconvolution from $I(t) = I_u(t) + I_d(t)$

The current outbreak size $Q(t) = E(t) + I(t)$

The curve $E(t)$, obtained by deconvolution of $I(t)$ is considered valid up to about 10 days before the present (1 mean incubation period of 5.6 days plus 2 standard deviations). We fit $E(t)$ using a time varying AR model with lag 1, and forecast to the present before summing with $I(t)$ to obtain a nowcast of the current epidemic size.

## Time series of new case reports $C(t)$ in China 

We extract a times series of new case reports $C(t)$ in China from the following table of new case reports by province:

<https://docs.google.com/spreadsheets/d/1gTj4zFMfXRk92CCQAOwa04ZBhqJ4oSmQFm_nb8-YXbw/edit#gid=1767307356>


```{r data, echo=TRUE, message = FALSE, warning = FALSE}

# read data directly to google sheet.

sheet <- "https://docs.google.com/spreadsheets/d/1gTj4zFMfXRk92CCQAOwa04ZBhqJ4oSmQFm_nb8-YXbw/edit#gid=1767307356"

china_province_data_cases_latest <- googlesheets4::sheets_read(sheet) 

# Create a Time Tibble; and sum provinces

fixcolumns <- function(x) {
  if(is.list(x)){
    is.na(x) <- lengths(x) == 0
    x <- gsub("[()]","",x)
    as.numeric(unlist(x))
  }
}

cases <- china_province_data_cases_latest %>% 
  rename_all(list(~make.names(.))) %>% 
  # select(-`Hubei (clinical)`) %>% 
  # rename(Date = `Date (CST)`) %>% 
  mutate_if(is.list, fixcolumns) %>% 
  mutate(Date = as.Date(Date)) %>% 
  padr::pad(start_val = as.Date("2019-12-01")) %>%
  replace(., is.na(.), 0) %>%
  tibbletime::as_tbl_time(index=Date) %>% 
  mutate(China = rowSums(.[,-1],na.rm = TRUE))

cases.china <- cases %>% select(Date, cases = China)

# plot
p_Cases <- plotly::plot_ly(data = cases, 
                x = ~Date , y = ~China,
                mode = 'lines+markers')
  

p_Cases_logy <- p_Cases %>% plotly::layout(yaxis = list(type = "log", range=c(-.25,5)))
  
  # plot(x=cases$Date,y=cases$China,type="o",pch=20,cex=.7,log='y')

p_Cases_logy
```


## Using deconvolution to backcast the number symptomatic and exposed.




If we know the incubation period, we can then reconstruct the time series of first exposure (class $E$ in and SEIR model) by a second deconvolution step. The incubation period of COVID-19 has been estimated.


### Deconvolution of time series of cases to recreate time series of symptom onset

We reconstruct a time series of symptom onset $I_s(t)$ from the time series of new case reports $C(t)$ using deconvolution. 

The beriod between symptom onset and case confirmation $\frac{1}{\gamma}$ is variable and changed throughout the early stages of the outbreak in Wuhan.

There are four different stages of the outbreak. Pre-January 9, January 9 to January 18, January 19 to January 23, and post-January 23. These timeframes are defined by Paige Miller in https://docs.google.com/document/d/1GpTvsoTZBWphn2tLbKHE5Z1ZE8RuD3caTHar-DqtqHc/edit?pli=1. The time between onset of symptoms and confirmation is distinct between these time frames. See <http://2019-coronavirus-tracker.com/stochastic.html> for more details.

**Table 1:** Comfirmation time (Period in days $\frac{1}{\gamma}$ between symptom onset and case confirmation). Mean, standard deviation and distribution of confirmation time for each stage of the outbreak.

```{r stages, echo=FALSE}

# Deconvolution Stages
stages <- tibble(
  stage = c(1:4),
  start = as.Date(c('2019-12-01','2020-01-09','2020-01-19','2020-01-24')),
  end = c(as.Date(c('2020-01-08','2020-01-18','2020-01-23')),max(cases$Date)),
  mean = c(6.57,4.77,.812,.2),
  sd = c(4.78,3.11,1.12,.45),
  distribution = "lognorm")
stages
```

First, we set up a single database to hold the original data, parameters, and all derived time series: 

```{r database, echo=TRUE, message = FALSE, warning = FALSE}
# set up container

stage <- list()
for (i in 1:nrow(stages)) {
  stage[[i]] <- create_series(stages$start[i] ~ stages$end[i], period = '1 d',  class = "Date") %>% 
  mutate(stage = i, mean = stages$mean[i], sd = stages$sd[i], distribution = stages$distribution[i])
}
database <- tbl_time(bind_rows(stage), index = date) %>% 
  mutate(cases = cases$China) %>% rename(Date = date)
remove(stage)
timespan <- length(database$Date)
```

We use a combination of the ridge regression transition matrix and random deconvolution to deconvolve the time-variant convolution. This method has not been tested yet, but it seems to work. 

```{r, echo=TRUE, message = FALSE, warning = FALSE}
estimates <- deconvolve_infection_curve_random_tv(as.data.frame(database))
incubation_matrix <- incubation_period_distribution_matrix_tv(
  outbreak_parms = as.data.frame(database), 
  rl = TRUE)
final_estimate <- deconvolve_infection_curve_rl(
  symptom_onset_curve = database$cases,
  matrix = incubation_matrix,
  estimate = estimates
  )
database$symptomatic <- round_infection_curve(database$cases, final_estimate)

# # plot
# p_Symptomatic <- plotly::plot_ly(data = database, x = ~Date , y = ~cases, name = 'confirmed', 
#                      type = 'scatter', mode = 'lines') %>% 
#   plotly::add_trace(y = ~symptomatic, name = 'symptom onset (estimated)', mode = 'lines')
#   


### Plot

col.cases <- 'rgba(0, 0, 0, .75)'
col.symptomatic <- 'rgba(230, 7, 7, .75)'
col.I_d <- 'rgba(230, 7, 7, .75)'
col.I_u <- 'rgba(230, 7, 7, .50)'
col.I <- 'rgba(230, 7, 7, .25)'
col.exposed <- 'rgba(7, 164, 181, 0.75)'
col.exposed.ci <- 'rgba(7, 164, 181, 0.10)'
col.E_d <- 'rgba(7, 164, 181, 0.75)'
col.E_u <- 'rgba(7, 164, 181, 0.50)'
col.E <- 'rgba(7, 230, 230, 1.0)' # brighter
col.fcast <- 'rgba(7, 7, 230, 0.75)'
col.fcast.ci <- 'rgba(7, 7, 230, 0.25)'

p_Symptomatic <- plotly::plot_ly(data = database, x = ~Date , y = ~cases, type = 'scatter',
                             name = 'newly confirmed', mode = 'lines',
                             line = list(color = col.cases)
                             ) %>% 
  plotly::add_trace(y = ~symptomatic, 
                    name = 'symptom onset (estimated)', mode = 'lines',
                    line = list(color = col.symptomatic, width=1))
  
p_Symptomatic %>% plotly::layout(yaxis = list(type = "log", range=c(-.25,5)))

```

### Time series of detected infectious individuals $I(t)$

$I_s(t)$ only gives us a time series of those *newly* infectious (detactable) cases. To find the number infectious (detectible) cases $I_d(t)$, we sum the number of newly infectious over the natural infectious period $\frac{1}{\gamma_0}$

```{r, echo=TRUE, message = FALSE, warning = FALSE}
database <- database %>% 
  mutate(I_d = zoo::rollapply(database$symptomatic, width = 7, FUN = sum, partial = TRUE, align = "right"))


p_nowcast <- plotly::plot_ly(data = database, x = ~Date , y = ~cases, type = 'scatter',
                             name = 'newly confirmed', mode = 'lines',
                             line = list(color = col.cases)
                             ) %>% 
  plotly::add_trace(y = ~symptomatic, 
                    name = 'symptom onset (detected)', mode = 'lines',
                    line = list(color = col.symptomatic, width=1)) %>% 
  plotly::add_trace(y = ~I_d, 
                    name = 'I (detected)', mode = 'lines',
                    line = list(color = col.I_d, width=2))

  
p_nowcast %>% plotly::layout(yaxis = list(type = "log", range=c(-.25,5)))

```

### Accounting for undetected Infectious cases

$I_d(t)$ represents the number of infectious whose cases were later confirmed. This curve includes neither individuals who were sympotmatic but whose cases were never reported (they were missed in the reporting process), nor the number of people infected but asympotmatic (people who never develop symptoms and therefore were also missed in the reporting process). We can simplify accounting for undetected cases by assuming only two subsets of cases: detected and undetected (asymptomatic or weakly symptomatic), and assume that all strongly symptomatic cases were detected.

In this case, undetected infectious cases depend on the detection probability $q(t)$ and coefficient of relative transmissibility $c$, which represents the transmissibility of undetected cases relative to detected cases.

$$
\text{let } Y = \text{total number of cases Infectious or Asymptomatic} \\
I_d = qY \\
Y = \frac{I_d}{q}\\
Y_u = Y - I_d \\
I_u = cY_u \\
I_u = c(Y - I_d) \\
I_u = c \left( \frac{I_d}{q} - I_d \right) \\
I_u = cI_d \left( \frac{1}{q} - 1 \right) \\
I = I_u + I_d
$$


```{r}
# Add basic parameters
database <- database %>% 
  mutate(q = q_fcn(Date)) %>% # Add q
  mutate(c = c_fcn()) %>% 
  mutate(a = a_fcn()) %>% 
  mutate(gamma0 = gamma0_fcn()) %>% # natural recovery/removal rate
  mutate(gamma = 1/mean) %>% # effective recovery/removal rate
  mutate(sigma = sigma_fcn()) # rate from E to I

# Add I_u and I
database <- database %>% 
  mutate(I_u = c * I_d * ((1/q)-1)) %>% 
  mutate(I = I_u + I_d)

p_nowcast <- plotly::plot_ly(data = database, x = ~Date , y = ~cases, type = 'scatter',
                             name = 'newly confirmed', mode = 'lines',
                             line = list(color = col.cases)
                             ) %>% 
  plotly::add_trace(y = ~symptomatic, 
                    name = 'symptom onset (detected)', mode = 'lines',
                    line = list(color = col.symptomatic, width=1)) %>% 
  plotly::add_trace(y = ~I_d, 
                    name = 'I (detected)', mode = 'lines',
                    line = list(color = col.I_d, width=2)) %>% 
  plotly::add_trace(y = ~I_u, 
                    name = 'I (undetected)', mode = 'lines',
                    line = list(color = col.I_u, width=2,dash="dot")) %>% 
  plotly::add_trace(y = ~I, 
                    name = 'I', mode = 'lines',
                    line = list(color = col.I, width=4))
  
p_nowcast %>% plotly::layout(yaxis = list(type = "log", range=c(-.25,5)))

```

Changing $q$ abruptly on January 10 has the effect of introducing a abrupt drop in $I$ on that date. Perhaps this is not realistic and a continuously varying function for $q$ should be considered.

<!-- If we know the reporting rate of sympotmatic cases $\rho$, we can adjust the symptom onset curve accordingly by multiplying by $1 - \rho$. -->

### Accounting for asymptomatic cases separately from $q$

If the detection probability $q$ reflects only the detection of symptomatic cases and not asymptomatic cases, then we would also need account for asymptomatic cases.

If detection rate $q$ does not reflect $a$, the proportion of cases asymptomatic, then 
the $q$ should me multiplied by $1-a$ to obtain a true detection rate.

This would give

$$
I_u = cI_d \left( \frac{1}{q(1-a)} - 1 \right)
$$



<!-- We currently do not account for underreporting of symptomatic cases.  If the level of underreporting of sympotmatic cases is known, we could compensate by adjusting the symptomatic onset curve accordingly. UPDATE: Actually, q accounts for this.-->

```{r}
database <- database %>% 
  mutate(I_u = c * I_d * ((1/(q * (1-a)))-1)) %>% 
  mutate(I = I_u + I_d)

p_nowcast <- plotly::plot_ly(data = database, x = ~Date , y = ~cases, type = 'scatter',
                             name = 'newly confirmed', mode = 'lines',
                             line = list(color = col.cases)
                             ) %>% 
  plotly::add_trace(y = ~symptomatic, 
                    name = 'symptom onset (detected)', mode = 'lines',
                    line = list(color = col.symptomatic, width=1)) %>% 
  plotly::add_trace(y = ~I_d, 
                    name = 'I (detected)', mode = 'lines',
                    line = list(color = col.I_d, width=2)) %>% 
  plotly::add_trace(y = ~I_u, 
                    name = 'I (undetected)', mode = 'lines',
                    line = list(color = col.I_u, width=2, dash="dot")) %>% 
  plotly::add_trace(y = ~I, 
                    name = 'I', mode = 'lines',
                    line = list(color = col.I, width=4))
  
p_nowcast %>% plotly::layout(yaxis = list(type = "log", range=c(-.25,5)))
```


### Deconvolution of symptom onset curve to estimate exposure curve

Now we have a rough estimate of the number of people infectious each day. 

Next, we perform a deconvolution of the symptom onset curve $I_s(t)$ to estimate the number of people exposed each day. We estimate the exposure curve 20 times and calculate a 95% confidence interval. 

Deconvolution produces a time series of equal length to the original time series. However, the incubation period introduces an edge effect at the end of the time series. The incubation period is a gamma distribution with mean 5.6 days and shape parameter 6.2.  

We therefore discard the last 10 days of the exposure curve before the present (1 incubation period plus 2 standard deviations).

```{r, echo=TRUE, message = FALSE, warning = FALSE}

# standard model inputs
incubation.params <- c(incubation.mean,incubation.sd)
# parms <- incubation.params

# # OLD STANRD MODEL INPUTS
# shape=6.2
# mean=5.6
# scale=mean/shape
# sd=sqrt(shape*scale^2)
# 
# parms <- c(mean,sd)
# distribution <- "gamma"

# number of days at end of exposed time series to throw out
na.tail <- round(incubation.mean+incubation.sd*2) 


# perform deconvolution, return list of raw results (tail of each estimate is invalid)
exposure <- deconvolve_single_curve(database$symptomatic, incubation.params)

# Replace tail of each estimate with NA's and store in database
database$exposure.mean <- tail_na(exposure$random,na.tail)
database$exposure.estimates <- as_tibble(tail_na(exposure$random_estimates,na.tail),
                                        .name_repair = 'universal'
                                        )

ci <- rowwise_confidence_intervals(database$exposure.estimates,interval=.95)
database$exposure.lower95 <- ci$lower
database$exposure.upper95 <- ci$upper

### plot

p_nowcast <- plotly::plot_ly(data = database, x = ~Date , y = ~cases, type = 'scatter',
                             name = 'newly confirmed', mode = 'lines',
                             line = list(color = col.cases)
                             ) %>% 
  plotly::add_trace(y = ~symptomatic, 
                    name = 'symptom onset (detected)', mode = 'lines',
                    line = list(color = col.symptomatic, width=1)) %>% 
  plotly::add_trace(y = ~exposure.mean, 
                    name = 'exposure (detected, mean)', mode = 'lines',
                    line = list(color = col.exposed, width=1)) %>% 
  plotly::add_ribbons(ymin = ~exposure.lower95, ymax = ~exposure.upper95,
                      name = 'exposure 95% confidence', mode='lines',
                      line = list(color = col.exposed, width=.5),
                      fillcolor = col.exposed.ci
                      )
  
p_nowcast %>% plotly::layout(yaxis = list(type = "log", range=c(-.25,5)))

```

### Time series of Exposed detected individuals $E_d(t)$

```{r}
database <- database %>% 
  mutate(E_d.mean = zoo::rollapply(database$exposure.mean, 
                                   width = incubation.mean, 
                                   FUN = sum, partial = TRUE, align = "right")) %>% 
  mutate(E_d.lower95 = zoo::rollapply(database$exposure.lower95, 
                                   width = incubation.mean, 
                                   FUN = sum, partial = TRUE, align = "right")) %>% 
  mutate(E_d.upper95 = zoo::rollapply(database$exposure.upper95, 
                                   width = incubation.mean, 
                                   FUN = sum, partial = TRUE, align = "right"))

p_nowcast <- plotly::plot_ly(data = database, x = ~Date , y = ~cases, type = 'scatter',
                             name = 'newly confirmed', mode = 'lines',
                             line = list(color = col.cases)
                             ) %>% 
  plotly::add_trace(y = ~symptomatic, 
                    name = 'symptom onset (detected)', mode = 'lines',
                    line = list(color = col.symptomatic, width=1)) %>% 
  plotly::add_trace(y = ~exposure.mean, 
                    name = 'exposure (detected, mean)', mode = 'lines',
                    line = list(color = col.exposed, width=1)) %>% 
  # plotly::add_ribbons(ymin = ~exposure.lower95, ymax = ~exposure.upper95,
  #                     name = 'exposure 95% confidence', mode='lines',
  #                     line = list(color = col.exposed, width=.5),
  #                     fillcolor = col.exposed.ci
  #                     ) %>% 
  plotly::add_trace(y = ~E_d.mean, 
                    name = 'E (detected, mean)', mode = 'lines',
                    line = list(color = col.exposed, width=2)) %>% 
  plotly::add_ribbons(ymin = ~E_d.lower95, ymax = ~E_d.upper95,
                      name = 'E (detected, 95% confidence)', mode='lines',
                      line = list(color = col.exposed, width=.5),
                      fillcolor = col.exposed.ci
                      )
  
p_nowcast %>% plotly::layout(yaxis = list(type = "log", range=c(-.25,5)))

```

### Accounting for undetected Exposed cases

Assuming again, that detection rate $q$ does not reflect the proportion of cases asymptomatic $a$, then the number of undetected Exposed cases $E_u$ is calculated as follows:

$$
E_u = E_d \left( \frac{1}{q(1-a)} - 1 \right)
$$


```{r}
database <- database %>% 
  mutate(E_u.mean = E_d.mean * ((1/(q * (1 - a))) - 1)) %>% 
  mutate(E.mean = E_u.mean + E_d.mean) %>% 
  mutate(E_u.lower95 = E_d.lower95 * ((1/(q * (1 - a))) - 1)) %>% 
  mutate(E.lower95 = E_u.lower95 + E_d.lower95) %>% 
  mutate(E_u.upper95 = E_d.upper95 * ((1/(q * (1 - a))) - 1)) %>% 
  mutate(E.upper95 = E_u.upper95 + E_d.upper95)

p_nowcast <- plotly::plot_ly(data = database, x = ~Date , y = ~cases, type = 'scatter',
                             name = 'newly confirmed', mode = 'lines',
                             line = list(color = col.cases)
                             ) %>% 
  plotly::add_trace(y = ~symptomatic, 
                    name = 'symptom onset (detected)', mode = 'lines',
                    line = list(color = col.symptomatic, width=1)) %>% 
  plotly::add_trace(y = ~exposure.mean, 
                    name = 'exposure (detected, mean)', mode = 'lines',
                    line = list(color = col.exposed, width=1)) %>% 
  plotly::add_trace(y = ~E_d.mean, 
                    name = 'E (detected, mean)', mode = 'lines',
                    line = list(color = col.E_d, width=2)) %>% 
  plotly::add_trace(y = ~E_u.mean, 
                    name = 'E (undetected, mean)', mode = 'lines',
                    line = list(color = col.E_u, width=2, dash="dot")) %>%
  plotly::add_trace(y = ~E.mean, 
                    name = 'E (mean)', mode = 'lines',
                    line = list(color = col.E, width=2)) %>%
  plotly::add_ribbons(ymin = ~E.lower95, ymax = ~E.upper95,
                      name = 'E 95% confidence', mode='lines',
                      line = list(color = col.exposed, width=.5),
                      fillcolor = col.exposed.ci
                      )
  
  


p_nowcast %>% plotly::layout(yaxis = list(type = "log", range=c(-.25,5)))
```

## Fit the exposed curve $E$ and forecast

```{r, echo=TRUE, message = FALSE, warning = FALSE}

lag <- 1

exposed.model.tvar <- tvReg::tvAR(trim_na(database$E.mean), 
                          p = lag,  # number of lags
                          type = "none",  # model does not contain intercept
                          est="ll", # "local linear" non parametric estimation method
                          tkernel = "Gaussian")

# forecast 10 days forward
exposed.forecast <- forecast::forecast(exposed.model.tvar$fitted,na.tail) 

database$E.fit <- exposed.model.tvar$fitted %>% pad_na(lag, front = TRUE) %>% pad_na(na.tail)

database$E.fcast.mean <- exposed.forecast$mean %>% pmax(0) %>% 
  pad_na(timespan - na.tail, front = TRUE)
database$E.fcast.upper95 <- exposed.forecast$upper[,'95%'] %>% pmax(0) %>%
  pad_na(timespan - na.tail, front = TRUE)
database$E.fcast.lower95 <- exposed.forecast$lower[,'95%'] %>% pmax(0) %>%
  pad_na(timespan - na.tail, front = TRUE)

### Plot

p_nowcast <- plotly::plot_ly(data = database, x = ~Date , y = ~cases, type = 'scatter',
                             name = 'confirmed', mode = 'lines',
                             line = list(color = col.cases)
                             ) %>% 
  plotly::add_trace(y = ~symptomatic, 
                    name = 'symptom onset (detected)', mode = 'lines',
                    line = list(color = col.symptomatic, width=1)) %>% 
  plotly::add_trace(y = ~exposure.mean, 
                    name = 'exposure (detected, mean)', mode = 'lines',
                    line = list(color = col.exposed, width=1)) %>% 
  plotly::add_trace(y = ~E.mean, 
                    name = 'E (mean)', mode = 'lines',
                    line = list(color = col.exposed, width = 2)) %>%  
  plotly::add_trace(y = ~E.fit, 
                    name = 'E (fitted)', mode = 'lines',
                    line = list(color = col.exposed, width = 1, dash = 'dot')) %>% 
  plotly::add_trace(y = ~E.fcast.mean, 
                    name = 'E (forecast mean)', mode = 'lines',
                    line = list(color = col.E, dash = 'dot')) %>% 
  plotly::add_ribbons(ymin = ~E.fcast.lower95, ymax = ~E.fcast.upper95,
                      name = 'E (forecast 95% confidence)', mode='lines',
                      line = list(color = col.E, width = .5),
                      fillcolor = col.exposed.ci
                      )
  
p_nowcast %>% plotly::layout(yaxis = list(type = "log", range=c(-.25,5)))
```


## Estimate outbreak size

The mean esimated outbreak size $Q(t)$ at each time $t$ is the sum of the Infectious curve $I(t)$ and either the mean estimated exposed curve $E(t)$ or mean forcested exposed curve.

For now, upper and lower 95% confidence for the esimated outbreak size at each time $t$ is simply the sum of $I$ and the upper and lower 95% confidence for the $E$, or the forecast of $E$ curve.

Do we need to handle confidence intervals differently?  Should we do residual resampling?

```{r, echo=TRUE, message = FALSE, warning = FALSE}

database <- database %>% 
  mutate(nowcast.mean = rowSums(
    dplyr::select(., I, E.mean, E.fcast.mean), na.rm = TRUE)
    ) %>% 
  mutate(nowcast.upper95 = rowSums(
    dplyr::select(., I, E.upper95, E.fcast.upper95), na.rm = TRUE)
    ) %>% 
  mutate(nowcast.lower95 = rowSums(
    dplyr::select(., I, E.lower95, E.fcast.lower95), na.rm = TRUE)
    )

# plot


p_nowcast <- plotly::plot_ly(data = database, x = ~Date , y = ~cases, type = 'scatter',
                             name = 'confirmed', mode = 'lines',
                             line = list(color = col.cases)
                             ) %>% 
  plotly::add_trace(y = ~I, 
                    name = 'I', mode = 'lines',
                    line = list(color = col.symptomatic, width = 2)) %>%
  plotly::add_trace(y = ~E.mean, 
                    name = 'E (mean)', mode = 'lines',
                    line = list(color = col.exposed, width = 2)) %>%  
  plotly::add_trace(y = ~E.fit, 
                    name = 'E (fitted)', mode = 'lines',
                    line = list(color = col.exposed, width = 1, dash = 'dot')) %>% 
  plotly::add_trace(y = ~E.fcast.mean, 
                    name = 'E (forecast mean)', mode = 'lines',
                    line = list(color = col.E, dash = 'dot')) %>% 
  plotly::add_ribbons(ymin = ~E.fcast.lower95, ymax = ~E.fcast.upper95,
                      name = 'E (forecast 95% confidence)', mode='lines',
                      line = list(color = col.E, width = .5),
                      fillcolor = col.exposed.ci) %>% 
  plotly::add_trace(y = ~nowcast.mean, 
                    name = 'nowcast (mean)', mode = 'lines',
                    line = list(color = col.fcast, width = 2, dash = 'dot')) %>% 
  plotly::add_ribbons(ymin = ~nowcast.lower95, ymax = ~nowcast.upper95,
                      name = 'nowcast (95% confidence)', mode='lines',
                      line = list(color = col.fcast, width = .5),
                      fillcolor = col.fcast.ci
                      )
  
p_nowcast_logy <- p_nowcast %>% plotly::layout(yaxis = list(type = "log", range=c(-.25,5)))
p_nowcast_logy

```

## Estimating the Effective Reproduction Number $R_{eff}$

We can estimate $R_{eff}$ by using the ODEs for the SEIR model, and estimating the slope of $E$ ($\frac{dE}{dt}$).

$$
\frac{dE}{dt} = \beta S I - \sigma E \\
R_{eff} = \frac{\beta S I}{\gamma} \\
\beta S I = \frac{dE}{dt}+ \sigma E  \\
R_{eff} = \frac{\frac{dE}{dt}+ \sigma E}{\gamma} \\
$$

To estimate $\frac{dE}{dt}$ we take a moving window average of the first difference of $E$.


```{r}

# Slope of E
slope.bw <- 7 # must be positive integer > 1
slope.samples <- as.list(seq(-slope.bw+1,slope.bw,1))

database$E.mean.diff <- c(NA,diff(database$E.mean))
database$E.lower95.diff <- c(NA,diff(database$E.lower95))
database$E.upper95.diff <- c(NA,diff(database$E.upper95))


database <- database %>% 
  mutate(E.mean.slope = zoo::rollapply(database$E.mean.diff, width = slope.samples, 
                                       FUN = mean, partial = FALSE, fill = NA)
         ) %>% 
  mutate(E.lower95.slope = zoo::rollapply(database$E.lower95.diff, width = slope.samples, 
                                       FUN = mean, partial = FALSE, fill = NA)
         ) %>% 
  mutate(E.upper95.slope = zoo::rollapply(database$E.upper95.diff, width = slope.samples, 
                                       FUN = mean, partial = FALSE, fill = NA)
         )

# R_eff calculations 

database <- database %>% 
  mutate(R_eff.mean = (E.mean.slope + sigma)/gamma) %>% 
  mutate(R_eff.lower95 = (E.lower95.slope + sigma)/gamma) %>% 
  mutate(R_eff.upper95 = (E.upper95.slope + sigma)/gamma)


### plot

p_Reff <- plotly::plot_ly(data = database, x = ~Date , y = ~R_eff.mean, type = 'scatter',
                             name = 'R effective', mode = 'lines',
                             line = list(color = 'black', width=2)
                             )  %>%
  plotly::layout(shapes = list(
                   list(type = "line", line = list(color = "pink", width=1, dash = "dot"), 
                        xref = "x", yref = "y",
                        x0 = ~min(Date), x1 = ~max(Date), y0 = 1, y1 = 1)
                   )
                 )

  # plotly::add_ribbons(ymin = ~R_eff.lower95, ymax = ~R_eff.upper95,
  #                     name = 'R_effective (95% confidence)', mode='lines',
  #                     line = list(color = 'lightgrey', width = .5),
  #                     fillcolor = 'lightgrey'
  #                     )

p_Reff_logy <- p_Reff %>% 
  plotly::layout(yaxis = list(type="log", range = c(-2, 2)),
                 shapes = list(
                   list(type = "line", line = list(color = "pink", width=1, dash = "dot"), 
                        xref = "x", yref = "y",
                        x0 = ~min(Date), x1 = ~max(Date), y0 = 1, y1 = 1)
                   )
                 )
p_nowcast_logy <- p_nowcast %>% plotly::layout(yaxis = list(type = "log", range=c(-.25,5)))

subplots <- list(p_Reff, p_nowcast_logy)
dash <- plotly::subplot(subplots, nrows = length(subplots), shareX = TRUE, titleY = TRUE) %>%
  plotly::layout(
      xaxis = list(
        spikethickness = 1,
        spikedash = "dot",
        spikecolor = "black",
        spikemode = "across+marker",
        spikesnap = "cursor"
      ),
      yaxis = list(spikethickness = 0)
    )
dash
  
```

TODO: Find out why R_eff going below zero.  Is the calc right?
